Quality of Service with Software Defined Network.

### 2017-03-13 report/paper.tex
altre correzioni

\textbf{Forwarding table:}
\begin{itemize}
 \item \texttt{match fields:} fields that compare packets in order to look for correspondences. They contain the in_port, the packet's header and others optional fields.
 \item \texttt{priority:} priority in case of rows with more than one correspondence;

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \item \texttt{counter:} this field was improved every time a correspondence with a packet was matched;
 \item \texttt{istructions:} a set of action that the switch add on \emph{Action-set} of packet or modification on pipeline flux;
 \item \texttt{timeouts:} this is the maximum time afther that the rule exired;
 %\item \texttt{cookie:} valori utilizzati per statistiche riguardo la singola voce, quindi non vengono considerati nell'elaborazione dei pacchetti dati;
 \item \texttt{flags:} this field is use to differentiate the management of the voice;
 \\
 \\
\end{itemize}

\textbf{Match Constraints}

these are theprincipal match method that we can use with OpenFlow v1.0:
\begin{itemize}
 \item \texttt{in_port}: the port of swith;
 \item \texttt{dl_src/dl_dst}: 802.3 MAC source and destination of packet;
 \item \texttt{dl_type}: 802.3 EtherType;
 \item \texttt{nw_tos}: IPv4 ToS;
 \item \texttt{nw_proto}: IPv4 Protocol;
 \item \texttt{nw_src/nw_dst}: IPv4 Source/Destination;
 \item \texttt{tp_src/tp_dst}: TCP/UDP Source/Destination Port;
 \\
 \\
\end{itemize}

\textbf{Action}

With the OpenFlow v1.0 in practice we can do simply things for example:
\begin{itemize}
 \item \texttt{Output}: set output port;
 \item \texttt{SetDLSrc/Dst}: modify packet MAC address;
 \item \texttt{SetNWSrc/Dst}: modify packet IPv4 address;
 \item \texttt{SetNWTos}
 \item \texttt{SetTPSrc/Dst}
\end{itemize}

From OpenFlow v1.1 on we can modify Time To Live and other things that allow us to do something more.

@@

\subsection{Related Work}
  The work presented in
"Efficient topology discovery in OpenFlow-based Software Defined Networks" by F. Pakzad \cite{farzaneh}
has been implemented to improve the controller speed and efficiency as it make it more reactive.
The introduction to the statistics power in a SDN network using the OpenFlow protocol has been found
in the paper called "Getting traffic statistics from network devices in an SDN environment using OpenFlow" by D.J. Hamad \& others. [ put note here]
For traffic analysis we haven't found any paper describing the characteristics of
some types of flows, so we used our PCs to sniff the traffic. Traffic analysis will be discussed in section.
  \#TODO stato dell arte e paper di riferimento  prova a darci un occhio adrea please.

\section{Needs and Goals}\label{sec:obb} %obbiettivi ed esigenze controllare
The main gool of this experience is to develop a small SDN network in a lab with real hardware
in orther to understand in a practical level how to implement a \textit{Quality of Service} in a SDN scenarios.
Our network have to manage traffic with a global purpose, in order to evolve according to different parameters.

Moreover the SDN network need to interact with world wide web efficently.
\subsection{Connectivity}\label{sec:connectivity}
First thing to do is provide connectivity, letting the controller know the connection of a host and avoiding packet to be lost while the first
add the computer to its graphes and try to add the rules to let the traffic go from the source to the destination. This has been done doing a
\textit{controlled flooding}: when a PacketIn event arrives at the controller, this look into the data structure that saves the nodes connected.
\\ If source and destination nodes are already found, it set the flow rules from the source node' switch to the destination node switch. To the latter,
a PacketOut message is prepared with inside the packet arrived inside the PacketIn event and with destination the port of the destination switch connected
to the destination node. Then the message is sent to the destination switch, so that the first packet is not lost.
If, instead, the destination node is not found, the controller sends to the switch that generated the PacketIn event a PacketOut message
with destination all the ports of the switch itself and with payload, the message that arrived within the PacketIn message.
To avoid the network to be filled of the only packet sent, we kept the information regarding the switches on which the packet that was flooded:
if a PacketIn message arrives at the controller and both the packet and the switch got a PacketOut message within 2 seconds, the controller doesn't
do anything, as this is a loop.
\\ So this approach allowed us to make the network unaware that it's not a legacy one and allowed connectivity to be provided.
%correggere
\subsection{Adaptability of the network}
An adaptable Network has to real-time reroute traffic in order to ensure a lot of demands.
A general efficient Network has to manage some tasks for example: avoid links congestion, reroute traffic when some malfunction accure on a link
and running some link path algorithm. 
In addition our network has to be adaptable in orther to ensure a QoS type of service.

To reach this gools we need to have a global knowledge on the network and we have to modify the routing table on every routerboard with a
global scope.
The performance of a single connection and entire network can ensure various type of metrics.
In order to ensure the traffic shaping based on network informations, there is a need of a structure,
automatically updatable, where the link statistics can be saved to allow the best path for a data stream. 
The best metric will be described below.
\#TODO MATTEO
\subsection{Exterior comunication}
Our network needed to be able to communicate with the external internet to be useful for a demo which demonstrate the capabilities of a SDN network.
The first thing we needed to accomplish was to make the controller aware of what switch and which port of the switch is connected to the external internet:
to do so, we allowed our controller to be able to receive and process in-line parameters. We provided two parametes to the controller, so that
it could recognize where the traffic to the internet need to be sent. Theese two parameters were the mac address of the
port of a switch that is connected to the link and the network address of our network, to let the controller know if a request from a host was meant to another host
within the network or the data stream was intended to the internet. We provided only one gateway, as the Routeboard had few ports available for the packet switching
and only four SDN-capable switches to be able to simulate a real network that was generating traffic.

\section{Instrument}\label{sec:instrument}
\subsection{Controller}
The controller that we use is POX, the software is opensource and it is write with Python, it is based on SDN and support OpenFlow.
The software use the version 2.7 of Python.
It's repository is at this link \url{https://github.com/noxrepo/pox} and a lot of documentation about the progect
is present at this link \url{https://openflow.stanford.edu/display/ONL/POX+Wiki}.
The only problem to use this controller is that it support only the first version of OpenFlow1.0.
\\
The controller run inside a VirtualMachine on our personal computer,
the VM is offered by SDN hub on this site: \url{http://sdnhub.org/tutorials/sdn-tutorial-vm/}.
This VM is a 64-bit Ubuntu 14.04 that has a number of SDN software and tools installed.
% \#TODO matteo da correggere e completare
\subsection{Router Board}
\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.30]{images/rboard.png}
  \caption{\emph{Experience scenario}}
  \label{fig:topo}
\end{figure}

The router board that we used is MikroTik miniROUTER RouterBOARD 450G. This Router Board virtualize an OpenFlow switch that support
OpenFlow v1.0\cite{routerboard_doc}.
\begin{itemize}
   \item CPU		AR7161 680MHz
   \item Memory		256MB DDR SDRAM onboard memory
   \item Data storage	512MB onboard NAND memory chip, microSD card slot (on reverse)
   \item Ethernet 	Three 10/100/1000 Mbit/s Gigabit Ethernet ports supporting Auto-MDI/X
   \item Serial port 	One DB9 RS232C asynchronous serial port
\end{itemize}
\textbf{firmware} the current version is: '' da cercare col comando: system routerboard print ''
\newline
The \textbf{Operating System} that they use is MikroTik RouterOS \cite{routeboard_software}.
\newline
One limit of this routerboard is that it has only 5 ethernet ports and this is a limit because one of these ports must be
reserved for the connection with the controller, consequently we have to connect both host and other routerboards using only 4 ports.

%correggere

\section{Results}\label{sec:results}
\subsection{Controller}
We developed a controller that can handle a mesh network, which has loops, that avoid
the traffic flooding. The controller stores the switch and the connection between them;
knows which hosts are connected to each switch port, with the hypothesys that only one
host is connected at each port, unless the port is the one connected to the internet.
With the use of the statistics, the controller get the status of the network and, if necessary,
tries to adjust the traffic and put in the link with the higher path loss the torrent traffic,
make the connection between two gaming host the one with less delay and tries to redistribute all the traffic in a way
such that all the nodes have an average throughput as lowest as possible.
\\
\newline
We have used some principal function released with POX controller:
\begin{itemize}
 \item \texttt{pox.openflow.discovery.launch()} this function runs an algorithm in order to discovery the network. Every time a new
 topological change occur this function returns an event on POX with which we can update our topological knowledge.
 It does not provide the host event. For host event\cite{pox}.
 \item \texttt{pox.openflow.spanning_tree.launch()} this function is necessary with a presence of loop on topology
 because with a spanning_tree algorithm it avoid the broadcast storm of message. When we set up \texttt{OFPP_FLOOD} like output port action in a table's
 switch row if some spanning_tree algorithm run the \texttt{OFPP_FLOOD} is only theoretical infact the algorithm close some port for flooding service
 on the switch in orther to avoid loop\cite{pox}.
\end{itemize}
\subsection{Real Lab Scenario}
The topological scenario in which we have set up experience is the next:
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/topo.png}
 \caption{\emph{Experience scenario}}
 \label{fig:topo}
\end{figure}

We have use 4 MikroTik Router board that were provided us in the Communication Networks Lab at DEI in University of Padua.
The 4 routerboard is connected each other in order to design a loop. Every sigle board connect two host and is connected with a legacy switch that provide
to logically connected with controller.

\subsection{Graph and Network structures}
  In order to achieve the realtime adaptability and statistics storage we have developed ad-hoc structure.
  We use a specific python library called NetworkX \cite{networkx} to create a graph in witch we can storage ad-hoc node that represent
  resplectively switch or host element.
  This topological structure is updated live thanks to connection event and link discover offered by POX's libraries:
  \begin{itemize}
   \item \texttt{_handle_LinkEvent}: from openflow.discovery an event of this type give us two type of information; a new link was added and a still present link was expired;
   \item \texttt{_handle_ConnectionUp}: from core.openflow this event join up when a new switch was added on the network and give us
   some information of the switch, in particular its ports configuration;
   \item \texttt{_handle_ConnectionUp}: from core.openflow like the event explained above comes up when a switch was disconnected from the network;
   \item \texttt{_handle_port_status}: from core.openflow this event occure when on a particular switch there is a change on its port, for example
   some host connect or the link was expired;\\
  \end{itemize}

  The link on our network besides have many parameters in witch we can save network performance. 
  In orther to update every link parameters we had to creat ad hoc function
  because there isn't some function constructed yet.

@@

The last traffic analysis we made was with a DOTA[explain what it is] web match.
We saw that for that type of gaming session, the packet size is about \# bytes and involves both TCP and UDP ports.
  \\
  \\
  For every networks performance we have tryed to develop a specific solution. The OpenFlow main goal is to centralize controll and software
  complexity so the routerboard can made only some easy function in orther to contain the hardware complexity and mantainly a good speed\cite{qos_paper}.
  Il che ha portato a dover utilizzare stratagemmi.
  \#TODO %riformulare
    \subsubsection{Delay}
Monitoring round-trip time provides important insights for network troubleshooting and traffic engineering.
The  common  monitoring  technique  is  to  actively  send  probe packets  from  selected  vantage  points  (hosts  or  middleboxes).
In traditional  networks, the control over the network routing is
limited,  making  it  impossible  to  monitor  every  selected  path. The  emerging  concept  of  Software  Defined  Networking  sim-plifies
network  control.  However,  OpenFlow,  the  common  SDN protocol,   does   not   support   RTT   monitoring   as   part   of   its
specification. In this experience we have tryed to emply an tecnic to escape this problem\cite{rtt}.
\newline The idea is to measure the time that occure from sending a packet from controller to a switch and the time in which the packet return from
an other switch.
\\
\newline First we have to discovery round trip time from controller to every single switch on our topology.
In order to do this on the controller we start a timer, we send a proper packet to the switch specifing the \texttt{OFPP_CONTROLLER - Send to the controller}
option and when the message return to the controller we stop the timer and calculare RTT for the switch.
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/rtt2.png}
 \caption{\emph{RTT between controller and OF switch}}
 \label{fig:topo}
\end{figure}
\newline
If we need to calculate the delay on a specific link, the rtt has to calculate with the headed switch of this link.
\\
\newline
Now we create a proper message with unique identifier number, we start a timer \texttt{TM1} and then send this message to a first switch specifing the output port.
Then when the packet comes to the second switch it identify the particular message and sends to the controller that stops it's dedicated timer \texttt{TM2}.
\newline This part required a particolar flow entry on the second switch in orther to send to the controller only some particulare packet.
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/rtt1.png}
 \caption{\emph{Pck from controller to itself following a particular way.}}
 \label{fig:topo}
\end{figure}
\\
\newline At this point we can calculate the delay of the link.
\\
\\
 delay = (TM2 - TM1) - RTT1/2 - RTT2/2
\\
\\
Teorethically this method must works but in practice there are a lot of delay introduce from computational concestion on the virtualized controller.
What we observe is a estimated delay like some seconds that is impossible, infact if we made some \texttt{ping test} from one host to another host
in this SDN network we realize delay around 3ms.
\\

@@

\begin{thebibliography}{99}
	\bibitem{pox} POX \url{https://openflow.stanford.edu/display/ONL/POX+Wiki}
	\bibitem{openflow} OpenFlow \url{http://flowgrammable.org/sdn/openflow/}
	\bibitem{pox_repo} POX's repository \url{https://github.com/noxrepo/pox}
	\bibitem{VM} Virtual Machine \url{http://sdnhub.org/tutorials/pox/}
	\bibitem{router_board} Router Board \url{https://www.senetic.it/product/RB450G}
	\bibitem{networkx} NetworkX python's library \url{https://networkx.github.io/}
	\bibitem{ONF} Open Networking Foundation \url{https://www.opennetworking.org/sdn-resources/technical-library}
	\bibitem{ONF_report} v1.5.1 OpenFlow report \url{https://www.opennetworking.org/sdn-resources/technical-library}
	\bibitem{routerboard_doc} RouterBOARD 450G MikroTik documentation \url{http://www.routerboard.sk/files/pdf/rb450g_manual.pdf}
	\bibitem{routeboard_software} RouterBOARD MikroTik Software \url{https://mikrotik.com/software}
	\bibitem{farzaneh}  Efficient Topology Discovery in Software Defined Networks F. Pakzad, M. Portmann, W. L. Tan and J. Indulska
	School of ITEE, The University of Queensland Brisbane, Australia \url{https://www.researchgate.net/publication/271508903_Efficient_Topology_Discovery_in_OpenFlow-based_Software_Defined_Networks}
	\bibitem{qos_paper} OpenNetMon: Network Monitoring in OpenFlow Software-Defined Networks Niels L.
	M. van Adrichem, Christian Doerr and Fernando A. Kuipers Network Architectures and Services, Delft University of Technology
	Mekelweg 4, 2628 CD Delft, The Netherlands \url{https://pdfs.semanticscholar.org/ea94/7a9334484ee4c3bb6bcfb4d1590aece1b178.pdf}
	\bibitem{rtt} Efficient Round-Trip Time Monitoring in OpenFlow Networks A. Atary A. Bremler-Barr Interdisciplinary Center Herzliya \url{http://www.deepness-lab.org/pubs/infocom2016_grami.pdf}

>>>>

\textbf{Forwarding table}
\begin{itemize}
 \item \texttt{match fields:} fields that compare packets in order to look for correspondences. They contain the in_port, the packet's header and others optional fields;
 \item \texttt{priority:} this field is used when there are rows with more than one correspondence;
 \item \texttt{counter:} this field was improved every time a correspondence with a packet is matched;
 \item \texttt{istruction:} a set of actions that the switch add on \emph{Action-set} of the matched packet or some instruction 
 in order to modify the pipeline flux;
 \item \texttt{timeout:} this is the maximum time afther which the rule expire;
 \item \texttt{flag:} with this field we can differenciate similar voices in order to managed multiple table with similar rules;
 \\
 \\
\end{itemize}

\textbf{Match Constraints}
these are the principal matching method that we can use with OpenFlow v1.0:
\begin{itemize}
 \item \texttt{in_port}: the port of swith;
 \item \texttt{dl_src/dl_dst}: 802.3 MAC source and destination of packet;
 \item \texttt{dl_type}: 802.3 EtherType;
 \item \texttt{nw_tos}: IPv4 ToS;
 \item \texttt{nw_proto}: IPv4 Protocol;
 \item \texttt{nw_src/nw_dst}: IPv4 Source/Destination addresses;
 \item \texttt{tp_src/tp_dst}: TCP/UDP Source/Destination ports;
 \\
 \\
\end{itemize}

\textbf{Action} With OpenFlow v1.0 we can do basic tasks in practice, the most important are for example:
\begin{itemize}
 \item \texttt{Output}: set output port;
 \item \texttt{SetDLSrc/Dst}: modify packet's MAC address;
 \item \texttt{SetNWSrc/Dst}: modify packet's IPv4 address;
 \item \texttt{SetNWTos}: change ToS fields of message;
 \item \texttt{SetTPSrc/Dst}
\end{itemize}

From OpenFlow v1.1 on we can modify Time To Live and other options that allow us to do something more complex.

@@

\subsection{Related Work}
  The work presented in
"Efficient topology discovery in OpenFlow-based Software Defined Networks" by F. Pakzad 
has been implemented to improve the controller speed and efficiency as it make it more reactive \cite{farzaneh}.
\newline The introduction to the statistics power in a SDN network using the OpenFlow protocol has been found
in the paper called "Getting traffic statistics from network devices in an SDN environment using OpenFlow" by D.J. Hamad \& others \cite{stat}.
\newline For traffic analysis we haven't found any paper describing the characteristics of
some types of flows, so we used our PCs to sniff the traffic. \newline 
Traffic analysis will be discussed in section \ref{subsec:traf}.

\section{Needs and Goals}\label{sec:obb}
The main purpose of this experience is to develop a small SDN network in a lab with physical hardware
in order to understand how to improve a Quality of Service in a SDN scenery with a real scenario.
Our network have to manage traffic with a global purpose, in order to evolve according to different parameters.

Moreover the SDN network have to efficiently interact with the world wide web.

\subsection{Connectivity}\label{sec:connectivity}
First thing to do is provide connectivity, letting the controller know the connection of a host and avoiding packet to be lost while the first
add the computer to its graphes and try to add the rules to let the traffic go from the source to the destination. This has been done doing a
\textit{controlled flooding}: when a PacketIn event arrives at the controller, this look into the data structure that saves the nodes connected.
\\ If source and destination nodes are already found, it set the flow rules from the source node' switch to the destination node switch. To the latter,
a PacketOut message is prepared with inside the packet arrived inside the PacketIn event and with destination the port of the destination switch connected
to the destination node. Then the message is sent to the destination switch, so that the first packet is not lost.
If, instead, the destination node is not found, the controller sends to the switch that generated the PacketIn event a PacketOut message
with destination all the ports of the switch itself and with payload, the message that arrived within the PacketIn message.
To avoid the network to be filled of the only packet sent, we kept the information regarding the switches on which the packet that was flooded:
if a PacketIn message arrives at the controller and both the packet and the switch got a PacketOut message within 2 seconds, the controller doesn't
do anything, as this is a loop.
\\ So this approach allowed us to make the network unaware that it's not a legacy one and allowed connectivity to be provided.

\subsection{Adaptability of the network}
An adaptable Network has to real-time reroute traffic in order to ensure a lot of demands.
A general efficient Network has to manage some tasks, for example avoiding links congestion, reroute traffic when some malfunction accure on a link
and running a link path algorithm. 
In addition our network has to be adaptable in order to ensure a QoS type of service.

To reach this gools we needed to have a global knowledge on the network and we had to modify the routing table on every routerboard with a
global purpose.
The performance of a single connection and entire network can ensure various type of metrics.
In order to ensure the traffic shaping based on network informations, there is a need of an automatically updatable structure,
where the link statistics can be saved to allow the best path for a data stream. 
The best metric will be described below.

\subsection{Exterior comunication}
Our network needed to be able to communicate with the external internet to be useful for a demo which demonstrate the capabilities of a SDN network.
The first thing we needed to accomplish was to make the controller aware of what switch and which port of the switch is connected to the external internet:
to do so, we allowed our controller to be able to receive and process in-line parameters. We provided two parametes to the controller, so that
it could recognize where the traffic to the internet need to be sent. Theese two parameters were the mac address of the
port of a switch that is connected to the link and the network address of our network, to let the controller know if a request from a host was meant to another host
within the network or the data stream was intended to the internet. We provided only one gateway, as the Routeboard had few ports available for the packet switching
and only four SDN-capable switches to be able to simulate a real network that was generating traffic.

\section{Instrument}\label{sec:instrument}
\subsection{Controller}
The controller that we use is POX, the software is opensource and it is written with Python, based on SDN and supports OpenFlow.
The software uses the version 2.7 of Python.
It's repository, together with lot of documentation regarding the project can be found at these links \url{https://github.com/noxrepo/pox}, \url{https://openflow.stanford.edu/display/ONL/POX+Wiki}.
The only problem in using this controller is that it supports only the first version of OpenFlow1.0.
\\
The controller runs inside a VirtualMachine on our personal computer and the VM is offered by SDN hub on this site: \url{http://sdnhub.org/tutorials/sdn-tutorial-vm/}.
This VM is a 64-bit Ubuntu 14.04 that has a number of SDN software and tools installed.

\subsection{Router Board}
\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.30]{images/rboard.png}
  \caption{\emph{Experience scenario}}
  \label{fig:topo}
\end{figure}

The router board that we used is MikroTik miniROUTER RouterBOARD 450G. This Router Board virtualizes an OpenFlow switch that supports
OpenFlow v1.0\cite{routerboard_doc}.
\begin{itemize}
   \item CPU		AR7161 680MHz
   \item Memory		256MB DDR SDRAM onboard memory
   \item Data storage	512MB onboard NAND memory chip, microSD card slot (on reverse)
   \item Ethernet 	Three 10/100/1000 Mbit/s Gigabit Ethernet ports supporting Auto-MDI/X
   \item Serial port 	One DB9 RS232C asynchronous serial port
\end{itemize}
%\textbf{firmware} the current version is: '' da cercare col comando: system routerboard print ''

The \textbf{Operating System} that they use is MikroTik RouterOS \cite{routeboard_software}.
\newline
One limit of this routerboard is that it has only 5 ethernet ports and this is a limit because one of these ports must be
reserved for the connection with the controller, that is why we have to connect both host and other routerboards using only 4 ports.

\section{Results}\label{sec:results}
\subsection{Controller}
We developed a controller that can handle a mesh network with loops, avoiding
packets flooding. The controller stores the switch and the connection between them and knows which hosts are connected to each switch port starting from the hypothesys that only one
host is connected to each port, unless the port is the one connected to the internet.
With the use of statistics, the controller gets the status of the network and, if necessary,
tries to adjust the traffic and to put in the link with the higher path loss the torrent traffic.
On the other hand it sets up the connection for two gaming hosts going through a link with less delay and tries to redistribute all the 
traffic in a way that all the nodes have an average throughput as lowest as possible.
\\
\newline
We have used some principal functions released with POX controller:
\begin{itemize}
 \item \texttt{pox.openflow.discovery.launch()} this function runs an algorithm in order to discover the network. Every time a new
 topological change occurs this function returns an event on POX with which we can update our topological knowledge.
 It does not provide the host event\cite{pox}.

 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \item \texttt{pox.openflow.spanning_tree.launch()} this function is necessary with a presence of loop on topology
 because with a spanning_tree algorithm it avoid the broadcast storm of message. When we set up \texttt{OFPP_FLOOD} like output port action in a table's
 switch row if some spanning_tree algorithm run the \texttt{OFPP_FLOOD} is only theoretical infact the algorithm close some port for flooding service
 on the switch in order to avoid loop\cite{pox}.
\end{itemize}
\subsection{Real Lab Scenario}
The topological scenario in which we have set up experience is the next:
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/topo.png}
 \caption{\emph{Experience scenario}}
 \label{fig:topo}
\end{figure}

We have use 4 MikroTik Routerboards that were provided us in the Communication's Network Lab at DEI in University of Padua.
The 4 routerboards are connected each other in order to design a loop. Every sigle board connects two host and is connected with a 
legacy switch which provides logically connection with controller.

\subsection{Graph and Network structures}
  In order to achieve the realtime adaptability and statistics storage we developed ad-hoc structure.
  We used a specific Python's library called NetworkX \cite{networkx}. NetworkX allow us to create a graph in witch we can storage proper nodes
  that represent resplectively switchs or hosts elements.
  This topological structure is updated live thanks to connection events and links discover offered by POX's libraries:
  \begin{itemize}
   \item \texttt{_handle_LinkEvent}: from openflow.discovery an event of this type give us two types of information; a new link is added and 
   a still present link is expired;
   \item \texttt{_handle_ConnectionUp}: from core.openflow this event join up when a new switch is added on the network and give us
   some information of the switch, in particular its ports configuration;
   \item \texttt{_handle_ConnectionUp}: from core.openflow like the event explained above comes up when a switch is disconnected from the network;
   \item \texttt{_handle_port_status}: from core.openflow this event occure when on a particular switch there is a change about its ports, for example
   some host is connected or the link is expired;\\
  \end{itemize}

  The link on our network besides has many parameters in witch we can save network's performance. 
  In order to update every links parameters we had to create ad hoc functionality
  because there isn't some function constructed yet.

@@

The last traffic analysis we made was with a DOTA[explain what it is] web match.
We saw that for that type of gaming session, the packet size is about \# bytes and involves both TCP and UDP ports.
  \\
  \\
  For every networks performance we have tryed to develop a specific solution. The OpenFlow main goal is to centralize controll and software
  complexity so the routerboard can made only some easy function in order to contain the hardware complexity and mantainly a good speed\cite{qos_paper}.
  Il che ha portato a dover utilizzare stratagemmi.
  \#TODO %riformulare
    \subsubsection{Delay}
Monitoring round-trip time provides important insights for network troubleshooting and traffic engineering.
The  common  monitoring  technique  is  to  actively  send  probe packets  from  selected  vantage  points  (hosts  or  middleboxes).
In traditional  networks, the control over the network routing is
limited,  making  it  impossible  to  monitor  every  selected  path. The  emerging  concept  of  Software  Defined  Networking  sim-plifies
network  control.  However,  OpenFlow,  the  common  SDN protocol,   does   not   support   RTT   monitoring   as   part   of   its
specification. In this experience we have tryed to emply an tecnic to escape this problem\cite{rtt}.
\newline The idea is to measure the time that occure from sending a packet from controller to a switch and the time in which the packet return from
an other switch.
\\
\newline First we have to discovery round trip time from controller to every single switch on our topology.
In order to do this on the controller we start a timer, we send a proper packet to the switch specifing the \texttt{OFPP_CONTROLLER - Send to the controller}
option and when the message return to the controller we stop the timer and calculare RTT for the switch.
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/rtt2.png}
 \caption{\emph{RTT between controller and OF switch}}
 \label{fig:topo}
\end{figure}
\newline
If we need to calculate the delay on a specific link, the rtt has to calculate with the headed switch of this link.
\\
\newline
Now we create a proper message with unique identifier number, we start a timer \texttt{TM1} and then send this message to a first switch specifing the output port.
Then when the packet comes to the second switch it identify the particular message and sends to the controller that stops it's dedicated timer \texttt{TM2}.
\newline This part required a particolar flow entry on the second switch in order to send to the controller only some particulare packet.
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/rtt1.png}
 \caption{\emph{Pck from controller to itself following a particular way.}}
 \label{fig:topo}
\end{figure}
\\
\newline At this point we can calculate the delay of the link.
\\
\\
 delay = (TM2 - TM1) - RTT1/2 - RTT2/2
\\
\\
Teorethically this method must works but in practice there are a lot of delay introduce from computational concestion on the virtualized controller.
What we observe is a estimated delay like some seconds that is impossible, infact if we made some \texttt{ping test} from one host to another host
in this SDN network we realize delay around 3ms.
\\

@@

\begin{thebibliography}{99}
	\bibitem{pox} POX \url{https://openflow.stanford.edu/display/ONL/POX+Wiki}
	\bibitem{openflow} OpenFlow \url{http://flowgrammable.org/sdn/openflow/}
	\bibitem{pox_repo} POX's repository \url{https://github.com/noxrepo/pox}
	\bibitem{VM} Virtual Machine \url{http://sdnhub.org/tutorials/pox/}
	\bibitem{router_board} Router Board \url{https://www.senetic.it/product/RB450G}
	\bibitem{networkx} NetworkX python's library \url{https://networkx.github.io/}
	\bibitem{ONF} Open Networking Foundation \url{https://www.opennetworking.org/sdn-resources/technical-library}
	\bibitem{ONF_report} v1.5.1 OpenFlow report \url{https://www.opennetworking.org/sdn-resources/technical-library}
	\bibitem{routerboard_doc} RouterBOARD 450G MikroTik documentation \url{http://www.routerboard.sk/files/pdf/rb450g_manual.pdf}
	\bibitem{routeboard_software} RouterBOARD MikroTik Software \url{https://mikrotik.com/software}
	\bibitem{farzaneh}  Efficient Topology Discovery in Software Defined Networks F. Pakzad, M. Portmann, W. L. Tan and J. Indulska
	School of ITEE, The University of Queensland Brisbane, Australia \url{https://www.researchgate.net/publication/271508903_Efficient_Topology_Discovery_in_OpenFlow-based_Software_Defined_Networks}
	\bibitem{qos_paper} OpenNetMon: Network Monitoring in OpenFlow Software-Defined Networks Niels L.
	M. van Adrichem, Christian Doerr and Fernando A. Kuipers Network Architectures and Services, Delft University of Technology
	Mekelweg 4, 2628 CD Delft, The Netherlands \url{https://pdfs.semanticscholar.org/ea94/7a9334484ee4c3bb6bcfb4d1590aece1b178.pdf}
	\bibitem{rtt} Efficient Round-Trip Time Monitoring in OpenFlow Networks A. Atary A. Bremler-Barr Interdisciplinary Center Herzliya \url{http://www.deepness-lab.org/pubs/infocom2016_grami.pdf}
	\bibitem{stat} Getting traffic statistics from network devices in an SDN environment using OpenFlow Diyar Jamal Hamad , Khirota Gorgees Yalda, and Ibrahim Tanner Okumus
	\url{http://itas2015.iitp.ru/pdf/1570195931.pdf};
